---
title: "Getting all up in Brent's biz - Getting to know you"
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_depth: 1
---

```{r setup, include=FALSE}
library(DBI)
con <- dbConnect(odbc::odbc(), .connection_string = "Driver={ODBC Driver 13 for SQL Server};server={brento.database.windows.net};
database={brentodb};
uid={datasci};
pwd={nZY0*51lG^};")

# cred creation

#in master
#CREATE LOGIN datasci WITH password='nZY0*51lG^';
# in db
#CREATE USER datasci FROM LOGIN datasci;
#EXEC sp_addrolemember 'db_datareader', 'datasci';
#alter user datasci with default_schema=anon
```

```{r message=FALSE, results='hide'}
library(tidyverse)
library(dbplyr)
```

# Navigating the DB
## Available tables
```{r}
odbc::dbListTables(con, schema_name="anon")
```

## Schema definitions
```{r}
(con %>% 
  tbl(in_schema("information_schema","columns")) %>% 
  select_all(str_to_lower) %>% 
  select(table_name, column_name, is_nullable, data_type) %>% 
  arrange(table_name, ordinal_position) %>% 
  collect() ->
   schema)
```

## Relationships
```{r}
schema %>% 
  select(table_name, column_name) %>% 
  group_by(column_name) %>% 
  mutate(matches=n())%>% 
  ungroup() %>% 
  filter(matches>1)  ->
  schema_matches

schema_matches %>% 
  spread(column_name, matches)
```

```{r}
library(igraph)
library(ggraph)

schema_matches %>% 
  inner_join(schema_matches,by="column_name") %>% 
  filter(table_name.x<table_name.y) %>%
  select(from=table_name.x, to=table_name.y) %>% 
  graph_from_data_frame() %>%
  ggraph(layout="linear", circular=TRUE) +
  geom_edge_arc(colour="grey") +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```

# People

## Mailchimp

```{r}
con %>% 
  tbl("mailchimp_members") ->
  mailchimp_members

mailchimp_members
```

### Domains
We probably have a lot of personal email accounts. Based on [mailchimp's 2015 analysis](https://blog.mailchimp.com/major-email-provider-trends-in-2015-gmail-takes-a-really-big-lead/), we might expect the most popular to be:

- gmail
- hotmail
- yahoo


```{r}
mailchimp_members %>% 
  group_by(email_domain_id) %>% 
  summarise(n=n()) %>% 
  collect() ->
  domains

domains %>%  
  top_n(10, n) %>% 
  arrange(desc(n))
```


```{r}
library(ggplot2)
domains %>% 
  filter(n<250) %>% 
  ggplot(aes(x=log(n))) +
  geom_density()
```

I'm going to provisionally set some boundaries between public email domains, large companies, and small companies based on number of signups.

```{r}
read_csv("../data/EmailProviders.csv") %>% 
  mutate(org_type="public") ->
  isps
domains %>% 
  left_join(isps) %>% 
  mutate(org_type= case_when(
    !is.na(org_type) ~ "public"
    ,n>20 ~ "large"
    ,TRUE ~ "small"
  )) ->
  domains

domains
```

### Signup source
This indicates where people signed up or were loaded from.
```{r}
mailchimp_members %>% 
  count(signup_source) %>% 
  collect() %>% 
  mutate(signup_sources=str_count(signup_source,", ")+1,
         all_signups=str_split(signup_source,", "),
         n=1) %>% 
  unnest(all_signups) %>% 
  mutate(all_signups=coalesce(all_signups,"Unknown source")) %>% 
  spread(all_signups, n,fill = 0) ->
  signup_sources

signup_sources
```

### I want to get
This indicates what mailing list campaigns folks would like to receive.

```{r}
mailchimp_members %>% 
  count(i_want_to_get) %>% 
  collect() %>% 
  mutate(i_want_to_gets=str_count(i_want_to_get,", ")+1,
         want_to_gets=str_split(i_want_to_get,", "),
         n=1) %>% 
  unnest(want_to_gets) %>% 
  mutate(want_to_gets=coalesce(want_to_gets,"Unknown")) %>% 
  spread(want_to_gets, n,fill = 0) ->
  want_to_gets

want_to_gets
```

### Member rating
This is a mailchimp defined feature based on how engaged subscribers are. It isn't foolproof as it relies on people allowing the tracking pixels but it's usually pretty good.

```{r}
mailchimp_members %>% 
  count(member_rating) %>% 
  mutate(prop=as.numeric(n)/sum(n)) %>% 
  arrange(member_rating)
```

Two thirds of the members have relatively low engagement rates based on their member ratings.

### Location

```{r message=FALSE}
library(ggmap)
mailchimp_members %>%
  collect() %>% 
  mutate(longitude=as.numeric(str_replace(longitude,"'","")),
         latitude=as.numeric(str_replace(latitude,"'",""))) %>% 
  {ggmap(get_stamenmap(bbox = c(left = -180, bottom = -80, right = 179.9999, top = 85), zoom = 3)) +
      geom_point(data=., aes(x=longitude, y=latitude), colour="red", size=.5) +
      theme_nothing()
  }
```

```{r}
mailchimp_members %>% 
  count(country_code) %>% 
  arrange(desc(n))
```

If using country code as a feature, we'll need to add an explicit missing and lump together the smaller labels.

## Wordpress users
```{r}
con %>% 
  tbl("users") ->
  wp_users

wp_users
```

### Mailchimp overlap

```{r}
mailchimp_members %>% 
  full_join(wp_users, by="email_user_id",suffix=c(".m",".w")) %>% 
  collect() %>% 
  mutate(wp_match_type= case_when(
    !is.na(email_domain_id.m)& !is.na(email_domain_id.w) ~ "Both"
    , !is.na(email_domain_id.m) ~ "Mailchimp"
    , !is.na(email_domain_id.w) ~ "Wordpress"
  )) %>%
  select(email_user_id, wp_match_type) ->
  user_types

user_types %>% 
  count(wp_match_type)
```

There are substantially more mailchimp subscribers than users who directly interact with the site in a signed up fashion. <5k are pure wordpress users. It will be interesting to see where profitability is. Assuming folks have to register on wordpress to purchase, then the Wordpress only bunch or the ones registered on both are likely to be the ones with purchases. The Wordpress only ones with purchases might be managers or people with authority to make purchases versus the engaged people who are the influencers/decision makers.

## Wordpress User Insight

```{r}
con %>% 
  tbl("usin_user_data") ->
  wp_users_insight

wp_users_insight
```

### Locations
```{r}
wp_users_insight %>% 
  count(country) %>% 
  arrange(desc(n))
```

Location data of wordpress users broadly matches the distribution of users in the mailing list.

### Engagement
```{r}
wp_users_insight %>% 
  count(sessions) %>% 
  collect() %>% 
  ggplot(aes(x=log(sessions), y=n), data=.)+
  geom_col()
```

The vast majority of registered users have single session behaviours.

```{r}
wp_users_insight %>% 
  count(sessions) %>% 
  collect() %>% 
  mutate(prop=as.numeric(n)/sum(n) ) %>% 
  arrange(desc(n),desc(sessions)) %>% 
  mutate(cum_prop=cumsum(as.numeric(n)/sum(n)))
```

85% of registered users have 10 sessions or less.

### Devices
```{r}
wp_users_insight %>% 
  count(browser, platform) %>% 
  collect() %>% 
  ungroup() %>% 
  mutate(n=scales::percent(as.numeric(n)/sum(n))) %>% 
  spread(platform, n, fill = "0%")
```

## Client contacts
```{r}
con %>% 
  tbl("client_contact") ->
  client_contact

client_contact
```

```{r}
client_contact %>% 
  distinct() %>% 
  left_join(mailchimp_members, by="email_domain_id", suffix=c(".c",".m")) %>% 
  group_by(client_id,email_domain_id) %>% 
  summarise(n=n_distinct(email_user_id)) %>% 
  arrange(desc(n))
```

It looks like there are a number of clients who transacted through public email domains. There are also clients without any mailchimp subscribers. When building feature, we will need to account for these facts.

```{r}
client_contact %>% 
  distinct() %>% 
  left_join(mailchimp_members, by="email_domain_id", suffix=c(".c",".m")) %>% 
  group_by(client_id,email_domain_id) %>% 
  summarise(n=n_distinct(email_user_id)) %>% 
  filter(n<1)
```

## Training purchasers
There isn't a high level purchasers table. The data is stored at the sale level instead in `XXX`.

** data not currently available **

## Students
In approximately 1/16th of training transactions - an address for a student has been provided.
```{r}
(con %>% 
  tbl("woocommerce_order_items_meta") ->
  students)
```

### Mailchimp overlap
```{r}
students %>% 
  left_join(mailchimp_members, by="email_user_id", suffix=c(".s",".m")) %>% 
  group_by(email_user_id) %>%
  summarise(n=n(), mailchimp=n_distinct(email_domain_id.m)) %>% 
  group_by(mailchimp) %>% 
  summarise(n=sum(n))
```

Approx. 20% of students don't subscribe.


# Training sales

## Posts
Training purchases partly end up in the `posts` table. The `posts` schema definition doesn't play nice with the odbc driver. As a result, a subset of columns needs to be imported instead of live connected. We should also restrict to a specific post_type to save extraneous data.
```{r}
con %>% 
  dbGetQuery("select 
            id as post_id,
            post_date,
            post_date_gmt,
            post_name,
            post_status
            comment_count
            from anon.posts 
            where post_type='shop_order'") ->
  posts

posts
```

## Sales

The sales data is very basic without being able to see the extra meta table we're waiting for.

```{r}
con %>% 
  tbl("woocommerce_order_items") ->
  training_sales

training_sales
```

```{r}
training_sales %>% 
  count(order_id) %>% 
  collect() %>% 
  ggplot(data=., aes(x=n)) +
  geom_density()
```

Most people order a single training item in one transaction.

```{r}
training_sales %>% 
  count(order_id) %>% 
  filter(n>1) %>% 
  arrange(desc(n))
```

```{r}
training_sales %>% 
  count(order_item_name) %>% 
  arrange(desc(n))
```

*Need to fix this regex to get first two words*
```{r}
training_sales %>% 
  collect() %>% 
  mutate(highlevel_training= str_extract(
    str_to_lower(order_item_name),
    boundary("word"))) %>% 
  count(highlevel_training) %>% 
  arrange(desc(n))
```

*Need to work out if the free training passes are sommething to be included as a desired outcome when building models*

## Sales over time

**Sales only starts in 2017?**

```{r}
posts %>% 
  collect() %>% 
  mutate(period=zoo::as.yearmon(post_date_gmt)) %>% 
  count(period) ->
  monthly_training_sales

monthly_training_sales %>% 
  ggplot(data=., aes(x=period, y=n)) +
  geom_line() + 
  geom_smooth()
```

# Consultancy sales

```{r}
con %>% 
  tbl("client_sales") ->
  client_sales

client_sales
```

## Exclusions

```{r}
client_sales %>% 
  filter(transaction_type=="Invoice" & amount > 0 & !is.na(product_service)) ->
  client_sales_clean
```


```{r}
client_sales %>% 
  filter(transaction_type!="Invoice" | amount < 0 | is.na(product_service)) 
```

## Products
```{r}
client_sales_clean %>% 
  count(product_service)
```

```{r}
client_sales_clean %>% 
  collect() %>% 
  mutate(product_type=case_when(
    str_detect(product_service,"Consulting") ~ "Consulting"
    ,str_detect(product_service,"Training") ~ "Training"
    ,TRUE ~ "Other"
  )) %>% 
  count(product_type)
```

*Should training be lumped into the training sales or are they special?* 

*Do the different consulting packages need to be split out?*

## Recurring clients
```{r}
client_sales_clean %>% 
  count(client_id) %>% 
  arrange(desc(n))
```

There are clients with ongoing work and we'll want to decide how to handle these.
```{r}
client_sales_clean %>% 
  filter(client_id==408)
```

## Sales over time

```{r}
client_sales_clean %>% 
  group_by(sale_date) %>% 
  summarise(n=n(), quantity=sum(quantity), sales=sum(amount)) ->
  daily_sales

daily_sales %>% 
  collect() %>% 
  mutate(period=zoo::as.yearmon(sale_date)) %>% 
  group_by(period) %>% 
  summarise(n=sum(n), quantity=sum(quantity), sales=sum(sales)) ->
  monthly_sales

monthly_sales %>% 
  ggplot(data=., aes(x=period, y=n)) +
  geom_line() + 
  geom_smooth()
```

```{r}
monthly_sales %>% 
  ggplot(data=., aes(x=period, y=sales)) +
  geom_line() + 
  geom_smooth()
```


```{r}
client_sales_clean %>%
  collect() %>% 
  mutate(product_type=case_when(
    str_detect(product_service,"Consulting") ~ "Consulting"
    ,str_detect(product_service,"Training") ~ "Training"
    ,TRUE ~ "Other"
  )) %>% 
  mutate(period=zoo::as.yearmon(sale_date)) %>% 
  group_by(period, product_type) %>% 
  summarise(n=n(), quantity=sum(quantity), sales=sum(amount)) ->
  product_sales

product_sales %>% 
  ggplot(data=., aes(x=period, y=n,
        group=product_type, colour=product_type)) +
  geom_line() + 
  geom_smooth()
```

# Newsletter engagement

*Need campaign reports*
*Need sent reports*

http://developer.mailchimp.com/documentation/mailchimp/reference/campaigns/#

## Opens
*Need campaign IDs*

```{r}
con %>% 
  tbl("mailchimp_opens") ->
  mailchimp_opens

mailchimp_opens
```

### Opens over time

```{r}
mailchimp_opens %>% 
  collect() %>% 
  mutate(period=zoo::as.yearmon(click_timestamp)) %>% 
  count(period) ->
  monthly_mailchimp_opens

monthly_mailchimp_opens %>% 
  ggplot(data=., aes(x=period, y=n)) +
  geom_line() + 
  geom_smooth()
```

## Clicks
*Need campaign IDs*

```{r}
con %>% 
  tbl("mailchimp_clicks") %>% 
  collect() ->
  mailchimp_clicks

mailchimp_clicks
```

### Clicks over time

```{r}
mailchimp_clicks %>% 
  mutate(period=zoo::as.yearmon(click_timestamp)) %>% 
  count(period) ->
  monthly_mailchimp_clicks

monthly_mailchimp_clicks %>% 
  ggplot(data=., aes(x=period, y=n)) +
  geom_line() + 
  geom_smooth()
```

### Click destinations
```{r}
mailchimp_clicks %>% 
  mutate(domain=urltools::domain(click_url)) %>% 
  count(domain) %>% 
  arrange(desc(n))
```

